---
title: "02_analysis"
author: "Sglatt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
if (!require("EGAnet")) {
  install.packages("EGAnet")
  require("EGAnet")
}
if (!require("psychTools")) {
  install.packages("psychTools")
  require("psychTools")
}
if (!require("dplyr")) {
  install.packages("dplyr")
  require("dplyr")
}

# Folder for output
if (!dir.exists("02_output")) {
  dir.create("02_output")
}
```

# Data
Both created in 01_data, but with different item versions (one with UG screener, one without)
```{r data}
merged_rm_cleaned_items <- read.csv("01_output/RM_items_clean_UG_R15_PTvCBT_ORP_PIONEER_2026_04_19.csv")
merged_rm_cleaned_items_x_UG <- read.csv("01_output/RM_items_R15_PTvCBT_ORP_PIONEER_2025_03_15.csv")

# remove dataset identifier and clinical demos
merged_rm_ega_vars <- merged_rm_cleaned_items %>%
  select(-dataset_id, -X, -starts_with("clinical_demo"))

merged_rm_x_UG_ega_vars <- merged_rm_cleaned_items_x_UG %>%
  select(-dataset_id, -X, -starts_with("clinical_demo"))
```

# Dictionary
Import data dictionary created in script 01
```{r dictionary}
ORP_dictionary <- read.csv("01_output/ORP_dictionary_2026_09_18.csv")

ORP_dictionary_labels <- ORP_dictionary %>%
  rename(
    items = variable # To match the EGA output 
  ) %>%
  select(items, label)

head(ORP_dictionary_labels)
```

# Unique variable analysis
```{r ega 1 items}
# merged_rm_cleaned_items_x_UG <- merged_rm_cleaned_items %>%
#  filter(dataset_id != "UG_dat")

# Rename variables to remove rm_
# colnames(merged_rm_x_UG_ega_vars) <- gsub("^rm_", "", colnames(merged_rm_x_UG_ega_vars))

# Unique variable analysis to see overlap
rm_x_UG_uva <- UVA(data = merged_rm_x_UG_ega_vars)
rm_x_UG_uva

# Look at items to keep & remove
rm_x_UG_uva$keep_remove
```

# EGA round 1
```{r ega 1}
# EGA
rm_x_UG_uva_ega <- EGA(
  data = rm_x_UG_uva$reduced_data,
  corr = "auto",
  model = "glasso",
  algorithm = "walktrap",
  uni.method = "louvain",
  plot.EGA = TRUE,
  verbose = TRUE
)

summary(rm_x_UG_uva_ega)
plot(rm_x_UG_uva_ega, layout = "circle")

# Network loadings
net.loads(rm_x_UG_uva_ega)

# Boostrap for stability
rm_x_UG_uva_boot <- bootEGA(
  data = rm_x_UG_uva$reduced_data,
  corr = "auto",
  model = "glasso",
  algorithm = "walktrap",
  uni.method = "louvain",
  iter = 1000,
  seed = 123456
)

rm_x_UG_uva_boot

# Compare median structure between empirical and boot
rm_x_UG_uva_compare <- compare.EGA.plots(
  rm_x_UG_uva_ega, rm_x_UG_uva_boot,
  labels = c("Empirical", "Bootstrap")
)

# Summary stats about bootstrapped version
summary(rm_x_UG_uva_boot)

# Item stability
rm_x_UG_uva_boot_dim <- dimensionStability(rm_x_UG_uva_boot)
rm_x_UG_uva_boot_dim

# Where are the items going in the times that it isn't going in its dimension
rm_x_UG_uva_boot_dim$item.stability$item.stability$all.dimensions
```

Make a csv with the items in each dimension with the item content
```{r ega 1 labels}
# Make dataframe with item + label + dimension
item_dim <- rm_x_UG_uva_ega$dim.variables

# Add description from data dictionary 
item_dim_labels <- item_dim %>%
  left_join(ORP_dictionary_labels, by = "items")

View(item_dim_labels)

# Save 
write.csv(
  item_dim_labels,
  file = paste0(
    "02_output/rm_x_UG_uva_dim_labels_",
    format(Sys.Date(), "%Y_%m_%d"),
    ".csv"
  ),
  row.names = TRUE
)
```

# EGA round 1 figures
```{r ega 1 figures, eval=FALSE}
# save figures 
pdf(paste0("02_output/rm_x_UG_uva_ega_plot_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
plot(rm_x_UG_uva_ega)
dev.off()

pdf(paste0("02_output/rm_x_UG_uva_ega_plot_circle_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 13, height = 10)
plot(rm_x_UG_uva_ega, layout = "circle")
dev.off()

pdf(paste0("02_output/rm_x_UG_uva_ega_comparison_plot_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
rm_x_UG_uva_compare <- compare.EGA.plots(rm_x_UG_uva_ega, rm_x_UG_uva_boot,
  labels = c("Empirical", "Bootstrap")
)
dev.off()

pdf(paste0("02_output/rm_x_UG_uva_ega_stability_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
dimensionStability(rm_x_UG_uva_boot)
dev.off()

```

# Remove items based on EGA round 1
```{r ega 2 items}
# Remove the item prefix for easier plotting
# colnames(merged_rm_cleaned_items_x_UG) <- gsub("^rm_", "", colnames(merged_rm_cleaned_items_x_UG))

# Remove UVA-identified items from above and those with poor function from initial EGA
merged_rm_x_UG_27_ega_vars <- merged_rm_cleaned_items_x_UG %>%
  select(
    -`rm_2`, -`rm_3`, -`rm_13`, -`rm_21`, -`rm_22`, -`rm_34`, -`rm_45`, -`rm_70`, # per UVA

    -`rm_9`, -`rm_11`, -`rm_20`, -`rm_23`, -`rm_52`, -`rm_19`,
    -`rm_35`, -`rm_54`, -`rm_37`, -`rm_46`, -`rm_53`, -`rm_51`, -`rm_66`, -`rm_67`, -`rm_36`, -`rm_39`, -`rm_72`,
    -`rm_12`, -`rm_24`, -`rm_25`, -`rm_26`, -`rm_44`, -`rm_47`, -`rm_30`, -`rm_38`, -`rm_43`, -`rm_27`, -`rm_48`, -`rm_33`, -`rm_41`, -`rm_56`,
    # poor function
    
    -starts_with("clinical_demo"), -dataset_id, -X
  )

```

# EGA round 2 
```{r ega 2}
# EGA
rm_x_UG_27_ega <- EGA(
  data = merged_rm_x_UG_27_ega_vars,
  corr = "auto",
  model = "glasso",
  algorithm = "walktrap",
  uni.method = "louvain",
  plot.EGA = TRUE,
  verbose = TRUE
)

summary(rm_x_UG_27_ega)
plot(rm_x_UG_27_ega, layout = "circle")

# Network loadings
net.loads(rm_x_UG_27_ega)

# Boostrap for stability
rm_x_UG_27_boot <- bootEGA(
  data = merged_rm_x_UG_27_ega_vars,
  corr = "auto",
  model = "glasso",
  algorithm = "walktrap",
  uni.method = "louvain",
  iter = 1000,
  seed = 123456
)


# Compare median structure between empirical and boot
rm_x_UG_27_compare <- compare.EGA.plots(
  rm_x_UG_27_ega, rm_x_UG_27_boot,
  labels = c("Empirical", "Bootstrap")
)

# Summary stats about bootstrapped version
summary(rm_x_UG_27_boot)

# Item stability
rm_x_UG_27_boot_dim <- dimensionStability(rm_x_UG_27_boot)
rm_x_UG_27_boot_dim

# Where are the items going in the times that it isn't going in its dimension
rm_x_UG_27_boot_dim$item.stability$item.stability$all.dimension
```

Make a csv with the items in each dimension with the item content
```{r ega 2 labels}
# Make dataframe with item + label + dimension
item_dim_2 <- rm_x_UG_27_ega$dim.variables

# Add description from data dictionary 
item_dim_2_labels <- item_dim_2 %>%
  left_join(ORP_dictionary_labels, by = "items")

View(item_dim_2_labels)

# Save 
write.csv(
  item_dim_2_labels,
  file = paste0(
    "02_output/rm_x_UG_items_removed_dim_labels_",
    format(Sys.Date(), "%Y_%m_%d"),
    ".csv"
  ),
  row.names = TRUE
)
```

# EGA round 2 figures
```{r ega 2 figures, eval=FALSE}
pdf(paste0("02_output/rm_x_UG_items_removed_ega_plot_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
plot(rm_x_UG_27_ega)
dev.off()

pdf(paste0("02_output/rm_x_UG_items_removed_ega_plot_circle_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 13, height = 10)
plot(rm_x_UG_27_ega, layout = "circle")
dev.off()

pdf(paste0("02_output/rm_x_UG_items_removed_ega_comparison_plot_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
rm_x_UG_27_compare <- compare.EGA.plots(rm_x_UG_27_ega, rm_x_UG_27_boot,
  labels = c("Empirical", "Bootstrap")
)
dev.off()

pdf(paste0("02_output/rm_x_UG_items_removed_ega_stability_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
dimensionStability(rm_x_UG_27_boot)
dev.off()
```

# (09/25) EGA round 3
Retaining some of the drinking/social items despite low performance because of theoretical/content importance and seeing what happens

```{r ega 3 items}
# Remove UVA-identified items from above and those with poor function from initial EGA; 
# BUT, 
# the ones that are commented and not crossed out are from social/drinking items that we theoretically want to keep in; 
# Social items include 20, 47, 56, 66, 67, 72, and drinking items include 36, 37, 38. 

merged_rm_x_UG_27_ega_vars_r3 <- merged_rm_cleaned_items_x_UG %>%
  select(
    -`rm_2`, -`rm_3`, -`rm_13`, -`rm_21`, -`rm_22`, -`rm_34`, -`rm_45`, -`rm_70`, # per UVA

    -`rm_9`, -`rm_11`, # -`rm_20`, 
    -`rm_23`, -`rm_52`, -`rm_19`,
    -`rm_35`, -`rm_54`, 
    # -`rm_37`, 
    -`rm_46`, -`rm_53`, -`rm_51`, 
    # -`rm_66`, -`rm_67`, -`rm_36`, 
    -`rm_39`,
    # -`rm_72`,
    -`rm_12`, -`rm_24`, -`rm_25`, -`rm_26`, -`rm_44`,
    # -`rm_47`, 
    -`rm_30`, 
    # -`rm_38`, 
    -`rm_43`, -`rm_27`, -`rm_48`, -`rm_33`, -`rm_41`, 
    # -`rm_56`,
    # poor function
    
    -starts_with("clinical_demo"), -dataset_id, -X
  )
```

# EGA round 3
```{r ega 3}
# EGA
rm_x_UG_27_ega_r3 <- EGA(
  data = merged_rm_x_UG_27_ega_vars_r3,
  corr = "auto",
  model = "glasso",
  algorithm = "walktrap",
  uni.method = "louvain",
  plot.EGA = TRUE,
  verbose = TRUE
)

summary(rm_x_UG_27_ega_r3)
plot(rm_x_UG_27_ega_r3, layout = "circle")

# Network loadings
net.loads(rm_x_UG_27_ega_r3)

# Boostrap for stability
rm_x_UG_27_boot_r3 <- bootEGA(
  data = merged_rm_x_UG_27_ega_vars_r3,
  corr = "auto",
  model = "glasso",
  algorithm = "walktrap",
  uni.method = "louvain",
  iter = 1000,
  seed = 123456
)


# Compare median structure between empirical and boot
rm_x_UG_27_compare_r3 <- compare.EGA.plots(
  rm_x_UG_27_ega_r3, rm_x_UG_27_boot_r3,
  labels = c("Empirical", "Bootstrap")
)

# Summary stats about bootstrapped version
summary(rm_x_UG_27_boot_r3)

# Item stability
rm_x_UG_27_boot_r3_dim <- dimensionStability(rm_x_UG_27_boot_r3)
rm_x_UG_27_boot_r3_dim

# Where are the items going in the times that it isn't going in its dimension
rm_x_UG_27_boot_r3_dim$item.stability$item.stability$all.dimension
```

Make a csv with the items in each dimension with the item content
```{r ega 2 labels}
# Make dataframe with item + label + dimension
item_dim_3 <- rm_x_UG_27_ega_r3$dim.variables

# Add description from data dictionary 
item_dim_3_labels <- item_dim_3 %>%
  left_join(ORP_dictionary_labels, by = "items")

View(item_dim_3_labels)

# Save 
write.csv(
  item_dim_3_labels,
  file = paste0(
    "02_output/rm_x_UG_r3_items_removed_dim_labels_",
    format(Sys.Date(), "%Y_%m_%d"),
    ".csv"
  ),
  row.names = TRUE
)
```

# EGA round 3 figures
```{r ega 3 figures, eval=FALSE}
pdf(paste0("02_output/rm_x_UG_r3_items_removed_ega_plot_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
plot(rm_x_UG_27_ega_r3)
dev.off()

pdf(paste0("02_output/rm_x_UG_r3_items_removed_ega_plot_circle_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 13, height = 10)
plot(rm_x_UG_27_ega_r3, layout = "circle")
dev.off()

pdf(paste0("02_output/rm_x_UG_r3_items_removed_ega_comparison_plot_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
rm_x_UG_27_compare_r3 <- compare.EGA.plots(rm_x_UG_27_ega_r3, rm_x_UG_27_boot_r3,
  labels = c("Empirical", "Bootstrap")
)
dev.off()

pdf(paste0("02_output/rm_x_UG__r3_items_removed_ega_stability_", format(Sys.Date(), "%Y_%m_%d"), ".pdf"), width = 10, height = 10)
dimensionStability(rm_x_UG_27_boot_r3)
dev.off()
```

